# Individual geoms

**Learning objectives:**

- 


## Introduction {-}

```{r setup, include=FALSE}
library(ggplot2)
```

This document explores the fundamental **geoms** in ggplot2, which serve as building blocks for visualizations. These geoms require **x** and **y** aesthetics and generally understand **color**, **size**, and sometimes **fill**.

## Fundamental Geoms {-}

### Area Plot {-}
```{r, eval=FALSE}
ggplot(mpg, aes(x, y)) + 
  geom_area(fill = "lightblue", color = "black") +
  ggtitle("Area Plot")
```

### Bar Chart {-}
```{r, eval=FALSE}
ggplot(df, aes(x, y)) + 
  geom_bar(stat = "identity", fill = "blue") +
  ggtitle("Bar Chart")
```

### Line Plot {-}
```{r, eval=FALSE}
ggplot(df, aes(x, y)) + 
  geom_line(color = "red") +
  ggtitle("Line Plot")
```

### Scatterplot {-}
```{r, eval=FALSE}
ggplot(df, aes(x, y)) + 
  geom_point(size = 3, color = "purple") +
  ggtitle("Scatterplot")
```

### Polygon Plot {-}
```{r, eval=FALSE}
ggplot(df, aes(x, y)) + 
  geom_polygon(fill = "lightgreen", color = "black") +
  ggtitle("Polygon Plot")
```

### Tile Plot {-}
```{r, eval=FALSE}
ggplot(df, aes(x, y)) + 
  geom_tile(fill = "orange") +
  ggtitle("Tile Plot")
```

### Text Annotation {-}
```{r, eval=FALSE}
ggplot(df, aes(x, y, label = label)) + 
  geom_text(size = 5, color = "black") +
  ggtitle("Text Plot")
```

## Exercises {-}

### {-}
1. What geoms would you use to draw each of the following plots?
- **Scatterplot**: `geom_point()`
- **Line Chart**: `geom_line()`
- **Histogram**: `geom_histogram()`
- **Bar Chart**: `geom_bar(stat = "identity")`
- **Pie Chart**: `geom_bar(width = 1, stat = "identity") + coord_polar(theta = "y")`

### {-}
2. Differences between `geom_path()` and `geom_polygon()`
- `geom_path()` connects points in the order they appear.
- `geom_polygon()` connects points and **closes** the shape by linking the last point back to the first.

### {-}
3. Differences between `geom_path()` and `geom_line()`
- `geom_line()` connects points in order of **x values**.
- `geom_path()` follows the exact order of points in the dataset.

### {-}
4. Low-level geoms used in composite geoms
- **`geom_smooth()`**: Uses `geom_line()` and `geom_ribbon()`.
- **`geom_boxplot()`**: Uses `geom_segment()` and `geom_rect()`.
- **`geom_violin()`**: Uses `geom_density_ridges()`.



## Meeting Videos {-}

### Cohort 1 {-}

`r knitr::include_url("https://www.youtube.com/embed/URL")`

<details>
<summary> Meeting chat log </summary>

```
LOG
```
</details>
